Improving the Replicability of Psychological Science Through Pedagogy

### 2017-05-17 paper/paper.Rmd
Experiment -> study

A summary of the original experiments chosen for replication is given in Table \ref{tab:summary}. Students chose findings from a wide variety of domains for replication. One cluster of studies included investigations of memory and visual attention [e.g., @xu2015], for example in applied contexts like remembering "deleted" files [@storm2015] or tracking transitions between locations on smartphone apps [@liverence2015]. A second cluster came from studies of social perception and social judgment, including judgments of faces and voices [@ko2015;@sofer2015], as well as studies of attributions of modesty, creativity, and expertise [@atir2015;@scopelliti2015;@proudfoot2015]. Other studies investigated increasing retirement savings by orienting participants to the future [@lewis2015], legacy priming [@zaval2015], and the effects of math anxiety on performance [@wang2015]. Details of individual replications are available in SOM.

@@

All study participants were recruited on AMT. Individual sample sizes are given in Table \ref{tab:summary}. Each sample was recruited independently, using the same account but a different title. All experiments were approved by the Stanford University Institutional Review Board under protocol \#23274, "Reproducibility of psychological science and instruction."

@@

Repeat administration of experiments on AMT has been raised as an issue in some prior work [e.g., @chandler2014], but tracking participation in specific paradigms is an open challenge. We did not ask participants whether they had participated in similar research previously, as we suspected that asking this sort of question would lead to a large number of inaccurate responses. In addition, half of the AMT population is estimated to change every six months [@stewart2015], and we expected that most of the studies in our sample that had been conducted on AMT had been performed at least a year previously (though see SOM for discussion of one particular paradigm that has been used more recently).

Determining appropriate sample sizes is a major challenge in replication research [@button2013;@simonsohn2015]. Post-hoc power analyses are problematic because of the likelihood of inflation of effect sizes due to the "winner's curse" [@hoenig2001;@button2013], but -- especially within the constraints of a limited budget -- it can be impossible to follow more conservative guidelines in all cases. For example, @simonsohn2015 recommends 2.5x the original sample, which can be feasible for small or under-powered original studies but may lead to impractical or unnecessary recommendations for experiments that were initially large and/or adequately powered.

We eventually used a hybrid, case-by-case selection criterion that attempted to maximize the success of the project while staying within the constraints of our budget.^[We initially were allocated \$1,500 but one student contributed personal research funds, leading to a total cost of $\approx$ \$1,700 for all studies. We attempted to set payment for our studies at approximately \$6/hour based on timing estimated from Pilot A sessions.] For all criteria, experimenters recruited an additional 5% of participants to ensure we still achieved the desired criterion if some participants skipped through or did not complete the experiment. In four cases, we powered the replication attempt to 80\% post-hoc power. In four cases, we used the original sample size. In one case, the original experiment was powered well above what would be required to find the effect of interest in the post-hoc analysis, so we chose a smaller sample size that still would yield high power. In two others, we reduced the number of trials as part of the adaptation to Mechanical Turk (in one of these, changing the sample size to compensate). In the last case, the original study used a multi-step procedure where the first step consisted of stimulus generation; we decreased the N in the first stage of the procedure and focused on achieving high power at the second stage.

@@

All experiments were coded in JavaScript, HTML, and CSS so as to be run in a standard web browser. All analyses were written using R, a free and open-source platform for statistical analysis. A schematic of our class workflow is shown in Figure 1.

@@

In addition to as-needed guidance on projects, students went through a process in which each student's work was reviewed several times by both the instructor (MCF) and the teaching assistants (ENS and RXDH). Each student collected two pilot samples and review was coordinated with these. The first, "Pilot A," consisted of a handful of non-naive participants (e.g., the experimenter, other students). The goal of this pilot was to ensure that all needed data were being logged by the experiment software and that analytic scripts for the confirmatory analyses functioned appropriately. After Pilot A was completed, the instructor or a TA critiqued and reviewed the student's experimental script, analytic code, and resulting output.

Once requested changes were made, the student conducted "Pilot B," using a handful of naive participants recruited from AMT. The goal of this pilot was to ensure that participants were able to complete the experiment and did not report any substantial issues with instructions or technical details of the experiment. (All students were instructed to give participants a way to leave free-form comments at the end of the experiment, prior to debriefing). At the conclusion of Pilot B, both the instructor and a TA reviewed the student's analytic code and its outputs on the data for all confirmatory analyses. The goal of this code review was to ensure that all planned analyses were specified for the key test selected in the original article, and that all relevant exclusion criteria, manipulation checks, etc. specified by the original authors were implemented correctly.

@@

Despite the general emphasis in meta-analytic and reproducibility efforts on aggregating a single effect size of interest [@osc2015;@lipsey2001], it was often challenging for us to identify a single key statistical test for each study. Usually a substantial number of statistical tests were conducted, often within one or several regression models with multiple specifications. Thus, we interpret "key statistical test" results with caution. Despite this interpretive difficulty, following previous work, we attempted to use a single statistical test as our target for experiment planning and analysis.

@@

After data analysis, for each experiment, the student carrying out each experiment and the instructors all gave independent "replication ratings" to assess the subjective success of the replication. Our three-point rating system was based on the theoretical support for the original finding, "None" (no support; 0\%), "Partial" (some support; 50\%), and "Full" (replication consistent with original interpretation; 100\%). We found that the ratings of the student carrying out the replication and the instructors were in agreement for `r sum(d$replicated_TA == d$replicated_student,na.rm=T)` out of the `r nrow(d)` replications.  The average rating given by instructors and students was `r 100 * round(mean(d$replicated_TA, na.rm=TRUE), digits = 2)`\% and `r 100 * round(mean(d$replicated_student, na.rm=TRUE), digits = 2)`\% respectively.

@@

In this paper, we reported a series of 11 student replications of previously-published experiments from the 2015 volume of *Psychological Science*. We had two interlocking goals: first, to create a proof of concept for pedagogical replications, and second, to provide evidence on the ability of students to replicate recent findings in a top journal. Importantly, rather than attempting to gauge the truth of particular effects, our project aimed to assess the challenges of replicating findings -- selected on grounds of feasibility and interest -- within the constraints of a course project. Despite this narrower scope, we believe that insights from our approach are applicable to the broader goal of assessing replicability at scale.

@@

One of the barriers to more reproducible work in psychology is pedagogical: Some of the tools (programmatic statistical analysis, version control) can feel inaccessible or intimidating to beginners. Indeed, our students, who came from a variety of disciplinary backgrounds, had relatively little experience with web experiments or literate, reproducible data analysis. Yet by the end of the course, all had gained sufficient proficiency with the suite of technical and conceptual tools necessary to carry out a study independently. This experience suggests that it is possible to convey key concepts in sufficient depth that -- within a supportive environment -- students can quickly learn to use a reproducible toolset.

From a purely statistical point of view, though, our results were underwhelming. Despite our relatively close adherence to the original experimental protocols and relatively large sample sizes, the modal outcome of our projects was partial replication. These partial replications often were cases in which some hint of the original pattern was observed but the key statistical test was not statistically significant and showed both smaller effect size and lower evidential value than the same test in the original. We invite readers to browse the narrative descriptions of individual replication attempts in our SOM to see the ways that patterns of experimental findings can differ from one another beyond the significance of a single test.

@@

Indeed, no experiment we attempted to replicate was pre-registered (though some were internal replications), and the general pattern of evidence across studies was often more important to the authors' conclusions than any particular test in any single study.  Almost every study conducted multiple statistical tests, often associated with several statistical models with differing specifications. And in some cases, there was no conventionally-reported and easily calculable effect size measures [e.g., for mixed- and random-effect models, cf. @bakeman2005]. This set of features -- which, in our experience, are endemic to published psychological work, including our own -- makes replication research within the conventional statistical paradigm extremely challenging.

>>>>

A summary of the original empirical studies chosen for replication is given in Table \ref{tab:summary}. Students chose findings from a wide variety of domains for replication. One cluster of studies included investigations of memory and visual attention [e.g., @xu2015], for example in applied contexts like remembering "deleted" files [@storm2015] or tracking transitions between locations on smartphone apps [@liverence2015]. A second cluster came from studies of social perception and social judgment, including judgments of faces and voices [@ko2015;@sofer2015], as well as studies of attributions of modesty, creativity, and expertise [@atir2015;@scopelliti2015;@proudfoot2015]. Other studies investigated increasing retirement savings by orienting participants to the future [@lewis2015], legacy priming [@zaval2015], and the effects of math anxiety on performance [@wang2015]. Details of individual replications are available in SOM.

@@

All study participants were recruited on AMT. Individual sample sizes are given in Table \ref{tab:summary}. Each sample was recruited independently, using the same account but a different title. All replications were approved by the Stanford University Institutional Review Board under protocol \#23274, "Reproducibility of psychological science and instruction."

@@

Repeat administration of empirical work on AMT has been raised as an issue in some prior work [e.g., @chandler2014], but tracking participation in specific paradigms is an open challenge. We did not ask participants whether they had participated in similar research previously, as we suspected that asking this sort of question would lead to a large number of inaccurate responses. In addition, half of the AMT population is estimated to change every six months [@stewart2015], and we expected that most of the studies in our sample that had been conducted on AMT had been performed at least a year previously (though see SOM for discussion of one particular paradigm that has been used more recently).

Determining appropriate sample sizes is a major challenge in replication research [@button2013;@simonsohn2015]. Post-hoc power analyses are problematic because of the likelihood of inflation of effect sizes due to the "winner's curse" [@hoenig2001;@button2013], but -- especially within the constraints of a limited budget -- it can be impossible to follow more conservative guidelines in all cases. For example, @simonsohn2015 recommends 2.5x the original sample, which can be feasible for small or under-powered original studies but may lead to impractical or unnecessary recommendations for studies that were initially large and/or adequately powered.

We eventually used a hybrid, case-by-case selection criterion that attempted to maximize the success of the project while staying within the constraints of our budget.^[We initially were allocated \$1,500 but one student contributed personal research funds, leading to a total cost of $\approx$ \$1,700 for all studies. We attempted to set payment for our studies at approximately \$6/hour based on timing estimated from Pilot A sessions.] For all criteria, experimenters recruited an additional 5% of participants to ensure we still achieved the desired criterion if some participants skipped through or did not complete the study. In four cases, we powered the replication attempt to 80\% post-hoc power. In four cases, we used the original sample size. In one case, the original study was powered well above what would be required to find the effect of interest in the post-hoc analysis, so we chose a smaller sample size that still would yield high power. In two others, we reduced the number of trials as part of the adaptation to Mechanical Turk (in one of these, changing the sample size to compensate). In the last case, the original study used a multi-step procedure where the first step consisted of stimulus generation; we decreased the N in the first stage of the procedure and focused on achieving high power at the second stage.

@@

All replication studies were coded in JavaScript, HTML, and CSS so as to be run in a standard web browser. All analyses were written using R, a free and open-source platform for statistical analysis. A schematic of our class workflow is shown in Figure 1.

@@

In addition to as-needed guidance on projects, students went through a process in which each student's work was reviewed several times by both the instructor (MCF) and the teaching assistants (ENS and RXDH). Each student collected two pilot samples and review was coordinated with these. The first, "Pilot A," consisted of a handful of non-naive participants (e.g., the experimenter, other students). The goal of this pilot was to ensure that all needed data were being accurately logged and that analytic scripts for the confirmatory analyses functioned appropriately. After Pilot A was completed, the instructor or a TA critiqued and reviewed the student's experimental script, analytic code, and resulting output.

Once requested changes were made, the student conducted "Pilot B," using a handful of naive participants recruited from AMT. The goal of this pilot was to ensure that participants were able to complete the study and did not report any substantial issues with instructions or technical details of the study (All students were instructed to give participants a way to leave free-form comments at the end of the study, prior to debriefing). At the conclusion of Pilot B, both the instructor and a TA reviewed the student's analytic code and its outputs on the data for all confirmatory analyses. The goal of this code review was to ensure that all planned analyses were specified for the key test selected in the original article, and that all relevant exclusion criteria, manipulation checks, etc. specified by the original authors were implemented correctly.

@@

Despite the general emphasis in meta-analytic and reproducibility efforts on aggregating a single effect size of interest [@osc2015;@lipsey2001], it was often challenging for us to identify a single key statistical test for each study. Usually a substantial number of statistical tests were conducted, often within one or several regression models with multiple specifications. Thus, we interpret "key statistical test" results with caution. Despite this interpretive difficulty, following previous work, we attempted to use a single statistical test as our target for replication planning and analysis.

@@

After data analysis, for each replication, the student carrying out each study and the instructors all gave independent "replication ratings" to assess the subjective success of the replication. Our three-point rating system was based on the theoretical support for the original finding, "None" (no support; 0\%), "Partial" (some support; 50\%), and "Full" (replication consistent with original interpretation; 100\%). We found that the ratings of the student carrying out the replication and the instructors were in agreement for `r sum(d$replicated_TA == d$replicated_student,na.rm=T)` out of the `r nrow(d)` replications.  The average rating given by instructors and students was `r 100 * round(mean(d$replicated_TA, na.rm=TRUE), digits = 2)`\% and `r 100 * round(mean(d$replicated_student, na.rm=TRUE), digits = 2)`\% respectively.

@@

In this paper, we reported a series of 11 student replications of previously-published empirical studies from the 2015 volume of *Psychological Science*. We had two interlocking goals: first, to create a proof of concept for pedagogical replications, and second, to provide evidence on the ability of students to replicate recent findings in a top journal. Importantly, rather than attempting to gauge the truth of particular effects, our project aimed to assess the challenges of replicating findings -- selected on grounds of feasibility and interest -- within the constraints of a course project. Despite this narrower scope, we believe that insights from our approach are applicable to the broader goal of assessing replicability at scale.

@@

One of the barriers to more reproducible work in psychology is pedagogical: Some of the tools (programmatic statistical analysis, version control) can feel inaccessible or intimidating to beginners. Indeed, our students, who came from a variety of disciplinary backgrounds, had relatively little experience with web-based empirical studies or literate, reproducible data analysis. Yet by the end of the course, all had gained sufficient proficiency with the suite of technical and conceptual tools necessary to carry out a study independently. This experience suggests that it is possible to convey key concepts in sufficient depth that -- within a supportive environment -- students can quickly learn to use a reproducible toolset.

From a purely statistical point of view, though, our results were underwhelming. Despite our relatively close adherence to the original protocols and relatively large sample sizes, the modal outcome of our projects was partial replication. These partial replications often were cases in which some hint of the original pattern was observed but the key statistical test was not statistically significant and showed both smaller effect size and lower evidential value than the same test in the original. We invite readers to browse the narrative descriptions of individual replication attempts in our SOM to see the ways that patterns of empirical findings can differ from one another beyond the significance of a single test.

@@

Indeed, no study we attempted to replicate was pre-registered (though some were internal replications), and the general pattern of evidence across studies was often more important to the authors' conclusions than any particular test in any single study.  Almost every study conducted multiple statistical tests, often associated with several statistical models with differing specifications. And in some cases, there was no conventionally-reported and easily calculable effect size measures [e.g., for mixed- and random-effect models, cf. @bakeman2005]. This set of features -- which, in our experience, are endemic to published psychological work, including our own -- makes replication research within the conventional statistical paradigm extremely challenging.
